A Hitchhiker's Guide to Reproducible Research" # replace with something serious

### 2021-01-26 manuscript.Rmd
explain some ways to simulate more realistic data

There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
A step to prevent unpleasant surprises is to simulate data.
Simulated data lets you verify that the data you expect to gather and the analysis that you plan fit together.
Of course one has to make a lot of educated guesses.
Generally there are of course hundreds of different equal possibilities, but we just care about the worst reasonable settings.
Sticking with our example about machiavilism, we simulate multivariate normal data representing the machiavilism subscale of the short dark triad (SD3).
As a starting point we can use already publishes information.
@Jones2013 report loadings seen in Table \@ref(tab:loadings) and report a gender difference (at the manifest level) depending on the sample between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.

@@

From which we can simulate data and fit a model to it:

```{r, results='hide'}
data <- lavaan::simulateData(model_truth, "cfa", sample.nobs = c(20000, 20000), meanstructure = TRUE, standardized = TRUE)
fit <- lavaan::cfa(model, data = data, group = "group", meanstructure = TRUE, std.lv = TRUE, group.equal = c("loadings", "intercepts")) 
lavaan::summary(fit)
```

<!--Hands-on: build simple functions tailored for your analysis-->

>>>>

There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
A step to prevent unpleasant surprises is to simulate data.
Simulated data lets you verify that the data you expect to gather and the analysis that you plan fit together.
Of course one has to make a lot of educated guesses to simulate data that closely matches your expectation.
These guesses strongly depend on the study you are planning and there cannot be a one fits all solution.
However, the analytic strategy often gives you a hint about how to simulate data.
In the machiavilism example the analytic strategy is to use structural equation modeling (SEM).
To simulate data for SEM one simply has to specify a model and then sample from a multivariate normal distribution following its implied covariance matrix.
The model parameters should mimic the machiavilism subscale of the short dark triad (SD3).
As a starting point we can use already publishes information.
@Jones2013 report loadings seen in Table \@ref(tab:loadings) and report a gender difference (at the manifest level) depending on the sample between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.

@@

From which we can simulate data:

```{r, results='hide'}
simulated <- lavaan::simulateData(
  model_truth,
  sample.nobs = c(1000, 1000),
  meanstructure = TRUE,
  standardized = TRUE
)
```

And then fit the statistical model:

```{r}
fit <-
  lavaan::cfa(
    model,
    data = simulated,
    group = "group",
    meanstructure = TRUE,
    std.lv = TRUE
  ) 
```

But wait `lavaan` warns us that this model may not be identified.
We overlooked a crucial assumption of this model, that is measurement invariance.
This assumption is also baked into our simulation of the data.
It is probably more realistic to violate this assumption and plan our analysis accordingly.
We therefore simulate the data so that some of the items parameter differ among groups and will then alter the analysis to assume partial measurement invariance.
While reconsidering our assumptions it is also unlikely that the data is normally distributed.
Not only would the real items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, therefore suggesting some skewness to the items.
@Rhemtulla2012 conclude that using Likert-like items is acceptable when they employ five or more categories.
We therefore just continue simulating continues data.
However, adding a little skewness to the items is easy and helps make the simulation more realistically.
Anyhow, remember that what we consider realistically for our case could be unfit for your case.
<!--Hands-on: build simple functions tailored for your analysis-->
